---
title: "Exercises for class 1"
output: html_document
date: "13-02-24" 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#install.packages(c("coda","mvtnorm","devtools","loo","dagitty","shape"))
#devtools::install_github("rmcelreath/rethinking")
```


```{r}
library(rethinking)
```



```{r}
# Install and load the rethinking package
install.packages("rethinking")
library(rethinking)


```

##Testing wether the R package Rethinking works properly (before the exercises)

```{r}
# Load a built-in dataset
data(Howell1)

# Explore the dataset
head(Howell1)


```

```{r}
fit <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * age,
    a ~ dnorm(178, 20),
    b ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ), data = Howell1
)

precise(fit)

```


```{r}
# Display model summary
precis(fit)
```

 

# Exercises for Class 1

## Supporting code

Below is code from Chapter 2 of the book that you can use to do the exercises. Not all the code you require is here though. Make sure to not rush it and take time to play around with each function. :))

```{r}
# Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20)

# compute likelihood at each value in grid
likelihood = dbinom(6, size = 9, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

## Exercises

### Easy

2E1. 

```{r}

# 2

```

2E2.

```{r}

# 3

```

2E3.

```{r}

# 1
```

2E4.

```{r}

# Discuss in pairs/groups. Ask me for a hint if you feel like it.
 
```

### Medium 

2M1.

```{r}

# (1) W,W W - Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20)
 
# compute likelihood at each value in grid
likelihood = dbinom(3, size = 3, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")
 
```

```{r}
# (2) W,W, W, L - Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20)
 
# compute likelihood at each value in grid
likelihood = dbinom(3, size = 4, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```


```{r}
# (3) L,W,W, L, W, W, W - Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20)
 
# compute likelihood at each value in grid
likelihood = dbinom(5, size = 7, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

2M2.

```{r}

# In the book, McElreath says that we can do much better than choosing a uniform prior for estimating the coverage of water. Is this what we are doing here? How does that impact our estimation?



# (1) W,W W - Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = ifelse(p_grid>0.5,1,0)
 
# compute likelihood at each value in grid
likelihood = dbinom(3, size = 3, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")


print(prior)
print(p_grid)
```


```{r}
# (2) W,W, W, L - Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = ifelse(p_grid>0.5,1,0)

# compute likelihood at each value in grid
likelihood = dbinom(3, size = 4, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")
```

```{r}
# (3) L,W,W, L, W, W, W - Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior 
prior = ifelse(p_grid>0.5,1,0)
 
# compute likelihood at each value in grid
likelihood = dbinom(5, size = 7, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points") 

```

2M3.

```{r}

# Use the Bayes formula!

# P(Earth, Land) = 0.15/0.65 = 0.23


     
```

For the exercises below, I highly suggest you to grab a piece of paper and try to solve the problems in a 'visual' way.

2M4.

```{r}

# Write your probability here.

```

2M5.

```{r}

# Write your probability here.

```

2M6.

```{r}

# Write your probability here.

```
